---
title: "Bitter and Temp survey data"
author: "Zane Chapman"
date: "2/11/2022"
output: html_document
---
load in library and packages

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(readr)
library(dplyr)
library(plotrix)
library(ggpmisc)
```

Upload raw survey data for both Tanner and RKC

```{r }
raw_data<-rbind(read_csv("data/05-07 Tanner RKC survey crab info.csv"), read_csv("data/08-10 Tanner RKC survey crab info.csv"),
                read_csv("data/11-13 Tanner RKC survey crab info.csv"), read_csv("data/14-15 Tanner RKC survey crab info.csv"),
                read_csv("data/16-18 Tanner RKC survey crab info.csv"),read_csv("data/19-21 Tanner RKC survey crab info.csv")
)

```

Select only Tanner crab and clean the names.  Summarise to get the total crab, then again to get the parasite information 

```{r }
raw_data<- raw_data %>% filter(Species=="Bairdi tanner crab")
raw_data<- clean_names(raw_data)

raw_data %>% 
  group_by(location, year, parasite) %>% 
  summarise(total_crab = sum(number_of_specimens))-> by_survey_area

by_survey_area %>% 
  spread(parasite, total_crab) %>% clean_names() -> parasite_sum
```

 Get the sum of all the parasite crab and clean crab to get a total amount of tanners
 
 Then clean to select only the relevant information 
```{r }
parasite_sum %>% 
  mutate(total_crab = sum( bitter_crab, none_present, briarosaccus_double_externa,
                          briarosaccus_double_scar, briarosaccus_single_externa, briarosaccus_single_scar,
                          carcinonemertes_errans, na, na.rm = T),
                          per_bitter = (bitter_crab/total_crab)*100, 
         per_bitter = round(ifelse(is.na(per_bitter), 0, per_bitter), 1)) -> bitter_summary
#clean up
bitter_summary<- bitter_summary %>% select(location, year, bitter_crab, total_crab, per_bitter)
```


Now we have a table with location, year, # of bitter crab, total # of crab and the percent of bitter crab for each location by year

We can create another table summary to show the bitter percent by year for each location.

```{r }
bitter_summary %>% 
  select(location, year, per_bitter) %>% 
  spread(year, per_bitter) %>% 
  arrange(desc(`2020`)) -> table_bitter
```

Now to input and create a table for the avg temperatures
```{r }
tidtemp<- read_csv("data/Tidbit master.csv") %>% clean_names() %>% select(year, location, depth_fathoms, meantemp)

```

clean and summarize to table to get the Standard Error of the average temp at each location and mean depth by year.
Note that there is a large difference in depths at which the temp is taken at, which is why the SE is included.
```{r }
tidtemp %>% group_by(location, year) %>% 
  summarise(SE= std.error(meantemp, na.rm = T), avgtemp= mean(meantemp, na.rm = T), mean.depth= mean(depth_fathoms, na.rm=T))  -> tidtemp_avg

```
merge the temp and bitter summary by location and year then remove the na

round the numbers to clean up and create a summary table of the the temperatures

```{r }
merge(tidtemp_avg, bitter_summary, by=c("location", "year")) %>% na.omit()  -> bitter_temp
bitter_temp$avgtemp<- round(bitter_temp$avgtemp, 2)
bitter_temp$SE<- round(bitter_temp$SE, 2)
bitter_temp$mean.depth<- round(bitter_temp$mean.depth, 2)

 #summary table for temps
 bitter_temp %>% 
    select(location, year, avgtemp) %>% 
    spread(year, avgtemp)  -> table_temp
```

Now we have object bitter_temp to use for exploratory analysis and plotting visuals


two plots for average temperature by year for each location.
SE is added to the figure to which for the most part is small
besides a few locations on specific years.

They were split into two figures to make it easier to view

```{r}
bitter_temp %>% filter(location=="Barlow Cove"|location=="Excursion Inlet"|location=="Glacier Bay"
                       |location=="Icy Strait"|location=="Juneau"|location=="Lynn Sisters"|
                         location=="Stephens Passage"|location=="Holkham Bay")%>% 
  ggplot(aes(year, avgtemp,  color=location))+
  geom_point()+
  geom_errorbar(aes(ymin= avgtemp-SE, ymax = avgtemp+SE), width = 0.2)+
  geom_line(aes(group=location))+
  xlab("Year")+ylab("Average Temperature")+labs(color="Location")

bitter_temp %>% filter(location=="Deadman Reach"|location=="Gambier Bay"|location=="Port Camden"|
                         location=="Port Frederick"|location=="Pybus Bay"|location=="Rodman Bay"|
                         location=="Seymour Canal"|location=="Thomas Bay")%>% 
  ggplot(aes(year, avgtemp,  color=location))+
  geom_point()+
  geom_errorbar(aes(ymin= avgtemp-SE, ymax = avgtemp+SE), width = 0.2)+
  geom_line(aes(group=location))+
  xlab("Year")+ylab("Average Temperature")+labs(color="Location")
```

Same thing for the bitter rate at each location

```{r}
bitter_temp %>% filter(location=="Barlow Cove"|location=="Excursion Inlet"|location=="Glacier Bay"
                       |location=="Icy Strait"|location=="Juneau"|location=="Lynn Sisters"|
                         location=="Stephens Passage"|location=="Holkham Bay")%>% 
  ggplot(aes(year, per_bitter,  color=location))+
  geom_point()+
  geom_line(aes(group=location))+
  xlab("Year")+ylab("Bitter Percent")+labs(color="Location")

bitter_temp %>% filter(location=="Deadman Reach"|location=="Gambier Bay"|location=="Port Camden"|
                         location=="Port Frederick"|location=="Pybus Bay"|location=="Rodman Bay"|
                         location=="Seymour Canal"|location=="Thomas Bay")%>% 
  ggplot(aes(year, per_bitter,  color=location))+
  geom_point()+
  geom_line(aes(group=location))+
  xlab("Year")+ylab("Bitter Percent")+labs(color="Location")
```
Here is where I start to look at the relationship of Bitter % and average temperature 
The first figure displays the bitter% ~ average temp

```{r}
ggplot(bitter_temp, aes(avgtemp, per_bitter))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) + geom_point()+
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = T)+ xlab("Average Temperature (C)")+ylab("Bitter Percent") 
```
I removed the temps greater than 7.5 since there are few and a very low bitter % associated with them

```{r}
bitter_temp %>% filter(avgtemp < 7.5) %>%  ggplot( aes(avgtemp, per_bitter))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) + geom_point()+
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = T)+ xlab("Average Temperature (C)")+ylab("Bitter Percent") 
```
Here I split the temperature from low (>10) and high (<10)

```{r}
bitter_temp %>% filter(per_bitter >10) %>%  ggplot( aes(avgtemp, per_bitter))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) + geom_point()+
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = T)+ xlab("Average Temperature (C)")+ylab("Bitter Percent") 
  
#plot for lower bitter %
  bitter_temp %>% filter(per_bitter <10, avgtemp<7.5) %>%  ggplot( aes(avgtemp, per_bitter))+
    geom_smooth(method = "lm", se=T, color="black", formula = y~x) + geom_point()+
    stat_poly_eq(formula = y~x, 
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                 parse = T) + xlab("Average Temperature (C)")+ylab("Bitter Percent")
```
After plotting these as a linear model there doesn't appear a strong relationship 
between temperature and bitter%.  It appears that location is a bigger factor for bitter % than temperature.

The next set of plots show bitter% as it relates to average temps with years shown as color points.

From these figures with a linear model most of the locations do show a negative relationship between
Bitter % and temperature, but the slopes and R^2 values differ between locations.

```{r}
bitter_temp %>% filter(location=="Barlow Cove") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+
  ggtitle("Barlow Cove")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")


bitter_temp %>% filter(location=="Deadman Reach") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Deadman Reach")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Excursion Inlet") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Excursion Inlet")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Gambier Bay") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Gambier Bay")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Glacier Bay") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Glacier Bay")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Holkham Bay") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Holkham Bay")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Icy Strait") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Icy Strait")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Juneau") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Juneau")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Lynn Sisters") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Lynn Sisters")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Port Camden") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Port Camden")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Port Frederick") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Port Frederick")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Pybus Bay") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Pybus Bay")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")

bitter_temp %>% filter(location=="Seymour Canal") %>%  ggplot( aes(avgtemp, per_bitter, color=year))+
  geom_smooth(method = "lm", se=T, color="black", formula = y~x) +
  stat_poly_eq(formula = y~x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) + geom_point()+ggtitle("Seymour Canal")+ xlab("Average Temperature (C)")+ylab("Bitter Percent")
```


